Tail Indices and Scale Parameters in Financial Time Series

### 2017-02-01 papers/FX/draft.tex
Some more work on the power utility function.

\subsection{power-law utility function}
In this section we follow Routledge and Zin \cite{routledge2010generalized} and assume
that the utility function of consumption $C$ is a power function%
\[
u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%
\]
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r$ be the return on the
riskless state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation*}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation*}

\subsubsection{mildly risk-avert utility function}
We first consider the class of utility functions for which $\gamma < 1$.
In this case $\E u(C) < \infty$ only if the distribution of $X$, call
it $F(\cdot)$, has a light upper tail. We assume this for the rest of
this section.

@@

Rewriting \eqref{eq:xxie0} gives
\begin{eqnarray*}
  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  && + \int_{\delta v}^{\infty}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  &=& I_1 + I_2 + I_3
\end{eqnarray*}
Only the first term $I_1$ involves the tail parameters $\alpha$
and $M$ in which we are interested, so we will only compute this
term. The integral of $I_1$ can be written as
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=& \int_{M}^{\infty}
[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
\end{eqnarray*}
Taylor expansion at $\phi e^{-x} = 0$ gives
\begin{eqnarray*}
  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
  &=& ((1 - \phi) e^r)^{1 - \gamma}
  + \sum_{n=1}^{\infty} {1 \over n!}
  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=&
{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
+ \sum_{n=1}^{\infty} {1 \over n!}
{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
\end{eqnarray*}
where $\Gamma(\alpha, x)$ is the upper incomplete gamma
function. For $x > 1$, this function may be expressed as (equation
8.11.2 of {\it NIST Digital Library of Mathematical Functions}
\cite{NIST.DLMF})
\[
\Gamma(a, z) = z^{a - 1} e^{-z}
\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
\]
Thus after re-arranging terms we get
\begin{eqnarray}
  {1 \over 1 + b} I_3
  &=&
  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
  &&
  + \sum_{k=0}^\infty
  {\alpha p \over M^{k+1} ((1 - \phi) e^r)^{\gamma - 1}}
  \prod_{i=0}^{k-1} (\alpha + i)
  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
  \left(
  {\phi \over e^{M + r} (1 - \phi)}
  \right)^n \times \nonumber \\
  && \prod_{i=1}^{n-1} (\gamma + i - 1)
  \label{eq:xxie1}
\end{eqnarray}
Observe that the sum over $n$ does not involve $\alpha$. In the
following we show this sum is positive for every $k \geq 0$, then it
is clear from \eqref{eq:xxie1} that $I_3$ and hence
$u(v) + b u(\delta v)$ is monotonically increasing with $\alpha$. To
see this, denote
\[
a_n(k) =
{(-1)^{n-1} \over n! n^{k+1}}
\left(
     {\phi \over e^{M + r} (1 - \phi)}
     \right)^n
     \prod_{i=1}^{n-1} (\gamma + i - 1)
\]
For each given $k$, $a_n(k)$ is an alternating series. To show the
positivity of $\sum_n a_n(k)$, it suffices to show
\[
  {a_n(k) \over |a_{n+1}(k)|} > 1
  \quad
  \text{ for } n=1,3,5,...
\]
Direct computation gives
\begin{eqnarray*}
  && {a_n(k) \over |a_{n+1}(k)|} \\
  &=& {
    e^{M + r} (1 - \phi) (n+1)^{k+2}
    \over
    \phi n^{k+1} (n + \kappa - 1)
  } \\
  &=&
  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
  \left( 1 + {1 \over n} \right)^{k+1}
  {e^{M + r} (1 - \phi) \over \phi}
\end{eqnarray*}

Thus we see that
$u(v) + u(\delta v)$ is monotonically increasing with $\alpha$. If
all investors are guided by and hence seek to maximize the same
utility function, they would all have constructed their portfolios by
choosing from a set of equities that have tail indices close to the
largest in the market. At market equilibrium, all actively traded
equities would have tail indices that are indistinguishable from each
other.

Notice that the summand of \eqref{eq:xxie1} decreases

>>>>

\subsection{power-law utility function}
In this section we follow Routledge and Zin \cite{routledge2010generalized} and assume
that the utility function of consumption $C$ is a power function%
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r$ be the return on the
riskless state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation*}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation*}

\subsubsection{mildly risk-avert utility functions}
We first consider the class of utility functions for which $\gamma < 1$.
In this case $\E u(C) < \infty$ only if the distribution of $X$, call
it $F(\cdot)$, has a light upper tail. We assume this for the rest of
this section.

@@

Rewriting \eqref{eq:xxie0} gives
\begin{eqnarray*}
  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  && + \int_{\delta v}^{\infty}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  &=& I_1 + I_2 + I_3
\end{eqnarray*}
Only the first term $I_1$ involves the tail parameters $\alpha$
and $M$ in which we are interested, so we will only compute this
term. The integral of $I_1$ can be written as
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=& \int_{M}^{\infty}
[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
\end{eqnarray*}
Taylor expansion at $\phi e^{-x} = 0$ gives
\begin{eqnarray*}
  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
  &=& ((1 - \phi) e^r)^{1 - \gamma}
  + \sum_{n=1}^{\infty} {1 \over n!}
  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=&
{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
+ \sum_{n=1}^{\infty} {1 \over n!}
{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
\end{eqnarray*}
where $\Gamma(\alpha, x)$ is the upper incomplete gamma
function. For $x > 1$, this function may be expressed as (equation
8.11.2 of {\it NIST Digital Library of Mathematical Functions}
\cite{NIST.DLMF})
\[
\Gamma(a, z) = z^{a - 1} e^{-z}
\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
\]
Thus after re-arranging terms we get
\begin{eqnarray}
  {1 \over 1 + b} I_3
  &=&
  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
  &&
  + \sum_{k=0}^\infty
  {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
  \prod_{i=0}^{k-1} (\alpha + i)
  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
  \left(
  {\phi \over e^{M + r} (1 - \phi)}
  \right)^n \times \nonumber \\
  && \prod_{i=1}^{n-1} (\gamma + i - 1)
  \label{eq:xxie1}
\end{eqnarray}
Observe that the sum over $n$ does not involve $\alpha$. Denote
\[
a_n(k) =
{(-1)^{n-1} \over n! n^{k+1}}
\left(
     {\phi \over e^{M + r} (1 - \phi)}
     \right)^n
     \prod_{i=1}^{n-1} (\gamma + i - 1)
\]
For each given $k$, $a_n(k)$ is an alternating series.
\begin{eqnarray*}
  && {a_n(k) \over |a_{n+1}(k)|} \\
  &=& {
    e^{M + r} (1 - \phi) (n+1)^{k+2}
    \over
    \phi n^{k+1} (n + \kappa - 1)
  } \\
  &=&
  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
  \left( 1 + {1 \over n} \right)^{k+1}
  {e^{M + r} (1 - \phi) \over \phi}
\end{eqnarray*}

If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_3$ and hence $u(v) +
u(\delta v)$ is monotonically increasing with $\alpha$. If all
investors are guided by and hence seek to maximize the same utility
function, they would all 
have constructed their portfolios by choosing from a set of equities
that have tail indices close to the largest in the market. At market
equilibrium, all actively traded equities would have tail indices that
are indistinguishable from each other.

\textcolor{red}{
  What happens when ${e^{M + r} (1 - \phi) \over \phi} \leq 1$?
}

\subsubsection{very risk-avert utility functions}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $ĸ =
\gamma - 1$.
